---
title: "Horizon 2020 work programmes"
subtitle: "Some text mining"
author: Jüri Lillemets
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output: 
  html_document:
    theme: yeti
    toc: true
editor_options:
  chunk_output_type: console
---
  
```{r setup, include = F}
# Settings
knitr::opts_chunk$set(include = T, eval = T, echo = F, 
                      message = F, warning = F, error = F, 
                      fig.height = 10, fig.width = 10, dpi = 75)
# Set working directory
setwd('/home/jrl/work/bioeastsup/horizon')
# Load packages
library('quanteda');library('quanteda.textmodels')
library('topicmodels')
library('wordcloud')
library('igraph');library('ggraph')
library('networkD3')
# Load data
load('corpus.Rda')
```

``` {r, include = F}
# Define some things
## Colors
ClPr <- c('#15436E', '#307108', '#3499DC', '#88C32C', '#4F81BD', '#4BACC6', '#DBEEF4')
#barplot(rep(1, length(ClPr)), names.arg = 1:length(ClPr), col = ClPr)
## Wordcloud function
plotWC <- function(words, freq, max = 200) {
  wordcloud::wordcloud(words, freq, #scale = c(5,.2), 
          random.order = F, 
          random.color = F, 
          max.words = max, min.freq = 1, 
          rot.per = 0, 
          fixed.asp = F,
          colors = colorRampPalette(c(ClPr[c(3,4,2)]))(10), 
          family = 'RobotoCondensed')
}
```

Activities were collected from the documents on the addresses below.

``` {r}
Urls <- c('https://ec.europa.eu/research/participants/data/ref/h2020/wp/2014_2015/main/h2020-wp1415-food_en.pdf', 
          'https://ec.europa.eu/research/participants/data/ref/h2020/wp/2016_2017/main/h2020-wp1617-food_en.pdf',
          'https://ec.europa.eu/research/participants/data/ref/h2020/wp/2018-2020/main/h2020-wp1820-food_en.pdf')
Urls
```

All words were stemmed, so that only stems of words and not the original words are shown (e.g. *sustainable* -> *sustain*). Stopwords were removed. Compounds were constructed from two or more words that often occurred together, here separated by an underscore (e.g. *sustainability in agriculture* -> *sustain* and *agricultur* -> *sustain_agriculture*).

#### Names of activities

``` {r}
names(Corp) %>% sub('^.*\\d{4}:\\s', '', .) # Remove activity code
```

A wordcloud from these names is below, based on word frequencies.

``` {r, fig.height = 8}
names(Corp) %>% sub('^.*\\d{4}:\\s', '', .) %>% plotWC
```

#### List of call topics

``` {r}
Calls <- c("Bio-based innovation for sustainable goods and services...", 
           "Blue Growth: Unlocking the potential of Seas and Oceans", 
           "Food and Natural Resources", 
           "Innovative, Sustainable and Inclusive Bioeconomy", 
           "Rural Renaissance - Fostering innovation and business opportunities",
           "Sustainable Food Security")
paste(levels(Corp$call), " - ", Calls)
```

#### Number of calls by years

``` {r}
table(Corp$call, Corp$year) %>% addmargins
```


# Term frequency

## Frequency of 50 most frequent terms

``` {r}
#head(textstat_frequency(dfMat))
topfeatures(dfMat, 50) %>% sort %>% dotchart(xlab = "Frequency")
```

## Wordcloud of all terms

``` {r}
freqs <- textstat_frequency(dfMat)
plotWC(freqs$feature, freqs$frequency)
```

## Wordcloud of unique terms

Only 200 most unique terms (based on document mean TF-IDF).

```{r}
tfIdf <- dfm_trim(dfMat, min_termfreq = 2) %>% 
  dfm_tfidf %>% as.matrix %>% apply(2, mean)
#plotWC(names(tfIdf), tfIdf)
uniTerm <- tfIdf %>% sort(decreasing = T) %>% names %>% head(200)
with(freqs[freqs$feature %in% uniTerm, ], plotWC(feature, frequency))
```

## Wordclouds of all terms by calls

```{r}
for (i in levels(Corp$call)) {
  print(Calls[match(i, levels(Corp$call))])
  dfm_subset(dfMat, Corp$call == i) %>% textstat_frequency %>% 
    with(plotWC(feature, frequency))
}
```

# Keyness

A positive value of keyness indicates that a term occurs more than expected in a target group compared to a reference group. Negative value indicates that it occurs less than expected. Here $\chi^{2}$ was chosen as the measure to estimate keyness.

Target is the group printed above plots and reference group contains all other documents. Ten terms with most negative and ten terms with most positive keyness are depicted on the plots below. Negative values are in blue, positie in green.

## By calls

``` {r}
for (i in levels(Corp$call)) {
  textstat_keyness(dfMat, target = Corp$call == i, measure = 'chi2') %>% 
    .[.$p < .05, ] %>% 
    #.[.$chi2 %>% abs %>% order(decreasing = T) %>% head(20), ] %>% .[order(.$chi2), ] %>% # First 20 by absolute values
    .[order(.$chi2), ] %>% with(rbind(head(., 10), tail(., 10))) %>% 
    #with(barplot(chi2, names.arg = feature, horiz = T, las = 2))
    with(dotchart(chi2, feature, #pt.cex = scale(n_target), 
         color = ClPr[c(3,2)][(chi2 > 0) + 1], 
         main = Calls[match(i, levels(Corp$call))]))
}
```

## By years

``` {r}
for (i in levels(Corp$year)) {
  textstat_keyness(dfMat, target = Corp$year == i, measure = 'chi2') %>% 
    .[.$p < .05, ] %>% 
    #.[.$chi2 %>% abs %>% order(decreasing = T) %>% head(20), ] %>% .[order(.$chi2), ] %>% # First 20 by absolute values
    .[order(.$chi2), ] %>% with(rbind(head(., 10), tail(., 10))) %>% 
    #with(barplot(chi2, names.arg = feature, horiz = T, las = 2))
    with(dotchart(chi2, feature, #pt.cex = scale(n_target), 
         color = ClPr[c(3,2)][(chi2 > 0) + 1],  
         main = i))
}
```

# Topic modelling

## Latent Dirichlet Allocation

This is an algorithm that attempts to find topics and returns a combination of words that would characterize each topic. 

The algorithm was ran so that it would find six topics (since there are 6 categories for calls). 

``` {r}
ldaMod <- LDA(dfMat, k = 6)
```

We can observe how do the six model predicted topics for each activity coincide with calls of each activity:

``` {r}
table(topics(ldaMod), Corp$call, dnn = c('predicted', 'actual'))
```

Below are six wordclouds describing each topic that LDA returned.

``` {r}
ldaTidy <- tidytext::tidy(ldaMod)
par(mfrow = 2:3, mar = rep(0, 4))
for (i in unique(ldaTidy$topic)) {
  with(ldaTidy[ldaTidy$topic == i & ldaTidy$beta > 1e-3, ], 
       plotWC(term, beta, max = 100))
}
```


# Term correlations

The network is based on Pearson's correlation between terms among documents.

``` {r, fig.height = 16}
corMat <- dfm_select(dfMat, max_nchar = 20) %>% dfm_trim(min_termfreq = 10) %>%
  as.matrix %>% cor
corMat[abs(corMat) < 0.6] <- 0
grMat <- graph_from_adjacency_matrix(corMat, mode = 'undirected', weighted = T, diag = F) %>% induced_subgraph(degree(.) > 5)
par(mar = rep(0,4))
plot(grMat, layout = layout_with_fr, 
     vertex.color = 'black', vertex.shape = 'circle', vertex.size = .5, 
     vertex.label.cex = .8, vertex.label.color = 'black', 
     vertex.label.family = 'RobotoCondensed', 
     edge.color = 'gray', edge.width = edge_attr(grMat)$weight %>% scale)
```

```{r, eval = F}
# Does not work well.
grD3 <- igraph_to_networkD3(grMat)
grD3$nodes$group <- 1
forceNetwork(grD3$links, grD3$nodes, NodeID = 'name', Group = 'group', 
             charge = -100, opacityNoHover = .8, opacity = 1, zoom = T)
```

```{r, eval = F}
# GGraph approach
m <- dfm_trim(dfMat, min_termfreq = 10) %>% as.matrix %>% cor
diag(m) <- 0
m <- data.frame(row=rownames(m)[row(m)], col=colnames(m)[col(m)], corr=c(m))
m <- m[m$corr > .65 & nchar(m$row) < 20 & nchar(m$col) < 20, ]
graph_from_data_frame(m) %>% 
  induced_subgraph(degree(.) > 5) %>% 
  ggraph(layout = 'fr') +
  geom_edge_link(aes(edge_alpha = corr), show.legend = F) +
  geom_node_point() +
  geom_node_text(aes(label = name, size = 6), show.legend = F, repel = T) +
  theme_graph()
```

```{r, echo = F}
m <- dfm_trim(dfMat, min_termfreq = 10) %>% as.matrix %>% cor
diag(m) <- 0
m <- data.frame(row=rownames(m)[row(m)], col=colnames(m)[col(m)], corr=c(m))
m <- m[m$corr > .6 & nchar(m$row) < 20 & nchar(m$col) < 20, ]
simpleNetwork(m, fontSize = 10, fontFamily = 'RobotoCondensed', 
                         linkColour = ClPr[3], nodeColour = 'black', 
                         opacity = 0.8, zoom = T) %>% 
  saveNetwork('network.html', selfcontained = TRUE)
```

An interactive network can be found [here](https://www.lillemets.ee/horizon/network).

# Text statistics

This section is not so relevant for investigating topics. But why not.

## Length of texts

``` {r}
Ngrams <- tokens_ngrams(Toks, n = 2:4)
par(mfrow = 1:2)
plot(nsentence(Corp) ~ Corp$call, xlab = 'Call', ylab = 'Sentences')
plot(nsentence(Corp) ~ Corp$year, xlab = 'Year', ylab = 'Sentences')
```

## Readability

Flesch's Reading Ease Score (Flesch 1948)^[Flesch, R. (1948). A New Readability Yardstick. Journal of Applied Psychology, 32(3), 221.] is based on number of words, syllabiles and average sentence length. Higher score indicates that a given text is easier to read.

``` {r}
Corp$readability <- textstat_readability(Corp)[, 2]
par(mfrow = 1:2)
boxplot(Corp$readability ~ Corp$call, xlab = 'Call', ylab = 'Readability')
boxplot(Corp$readability ~ Corp$year, xlab = 'Year', ylab = 'Readability')
```

## Lexical diversity

Carroll's Corrected Type-Token Ratio.

``` {r}
lexDiv <- textstat_lexdiv(dfMat, measure = 'CTTR')[[2]]
par(mfrow = 1:2)
boxplot(lexDiv ~ Corp$call, ylim = c(0, max(lexDiv)), las = 2, 
        xlab = 'Call', ylab = 'CTTR')
boxplot(lexDiv ~ Corp$year, ylim = c(0, max(lexDiv)), las = 2, 
        xlab = 'Year', ylab = 'CTTR')
```

## Sentiment

Sentiment was determined using 2015 Lexicoder Sentiment Dictionary.

``` {r}
Sent <- tokens_lookup(Toks, dictionary =  data_dictionary_LSD2015[1:2]) %>% 
  lapply(table) %>% do.call(rbind, .) %>% as.data.frame
```

### By call

```{r}
par(mfrow = 1:2)
for (i in c('negative', 'positive')) {
  boxplot(Sent[, i]/sapply(Toks, length) ~ Corp$call, 
          xlab = 'Call', ylab = 'Words with the given sentiment, %', main = i)
}
```

### By year

```{r}
par(mfrow = 1:2)
for (i in c('negative', 'positive')) {
  boxplot(Sent[, i]/sapply(Toks, length) ~ Corp$year, 
          xlab = 'Year', ylab = 'Words with the given sentiment, %', main = i, 
          las = 2)
}
```


## Correspondence analysis

Position of documents on a single dimension scale.

``` {r, fig.height = 16}
## Correspondence analysis
caMod <- textmodel_ca(dfMat)
caMod$rownames <- sub(':.*$', '', names(Corp))
textplot_scale1d(caMod, groups = Corp$call)
```
<!--
## Position of words

Position of word "sustain" in documents.

``` {r}
Toks[Corp$call == 'SFS'] %>% kwic(pattern = 'sustain') %>% textplot_xray
```
-->
